Supporting Information for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-02-20 writing/ehp_discussion.tex
Edit the text

Epidemiological studies can help characterize which health risks are elevated
during disasters, to what degree, and to
whom~\parencite{ibrahim2005unfortunate, noji2005disasters}.  As a result, these
studies help improve disaster preparedness and
response~\parencite{noji2005disasters}.  However, tropical cyclones are
multi-hazard events, making it complicated to measure exposure and so to
conduct multi-community, multi-year studies leveraging large administrative
datasets.  Here, we provide open-source county-level data for several tropical
cyclone exposures and explore limitations in that data.  
Further, we explore patterns in storm exposure classifications based on
different metrics, and we find that county-level tropical cyclone exposure
assessments vary substantially when using different metrics, based either on
storm hazards or on how close the storm's central track came to the community.
Our results can inform exposure assessment for future county-level studies of
the health risks and impacts associated with tropical cyclones exposure as
well as insights to inform epidemiological study design.

@@

The open-source data generally correspond well with data from other potential
sources (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison}
and~S2), but there are some caveats. The rainfall data are generally
well-correlated with ground-based observations, but may sometimes oversmooth
very high rainfall values (Figure~\ref{fig:raincomparison}). When rainfall data
is used to create binary exposure classifications, this disagreement is
unlikely to influence results, as both data sources agree in identifying these
as storms with high rainfall, but would be important to consider for cases that
include rainfall as a continuous measurement.

The peak sustained wind estimates are based on modeled, rather than observed,
values, and while the modeled wind data generally agree well with post-analysis
maximum wind radii from \ac{HURDAT2}~\parencite{landsea2013}
(Figure~\ref{fig:windcomparison}), there were a few storms with some
discrepancies. These storms---for example, Hurricane Sandy in 2012 and
Hurricane Ike in 2008---were unusually large systems for which high winds
persisted well inland from landfall (Figures~\ref{fig:windcomparison} and~S2).

For the flooding data, we found that flood event status, as determined based on
the NOAA Storm Events listings, typically agreed with measurements from
\ac{USGS} streamgages, with a flood event more likely to be listed if a storm
elevated streamflow at streamgages across the county
(Figure~\ref{fig:floodcomparison}).  However, there are differences between the
two flooding datasets, and these highlight both the difficulty of measuring
flood exposure at the county level and inherent challenges in using data from a
storm event database for epidemiologic exposure assessment.  For example, there
was one storm in Fulton County, GA, for which there was high streamflow but not
an associated flood event listing (Hurricane Ida, 2009).  This storm occurred
in November 2009, following a month with historic rainfall and flooding in
Georgia~\parencite{shepherd2011overview}.  In this case, the flooding
associated with Ida was incorporated into an ongoing flood event listing, with
a start date well before the five-day window we used to temporally match storm
event listings with tropical cyclone tracks.

This disagreement highlights the difficulty of large-scale pairing of storm
tracks with storm event listings---without criteria for temporally matching
event start dates to storm dates, many false positives would be
captured, for which the occurrance of a storm in the midst of an ongoing storm
event might be improperly attributed to the storm.  However, distance
and time restrictions, like those we used in matching NOAA Storm Event listings
with tropical cyclone locations and dates, do create the opportunity for
occasional false negatives, as for Hurricane Ida in Fulton County, GA, where a
storm contributes meaningfully to an ongoing event, but the event is not
captured for the storm in the exposure data.

@@

Tropical cyclone wind exposures had a strong coastal pattern, which is
consistent with the dramatic decrease in wind intensity that typically
characterizes the landfall of tropical cyclones.  Tropical cyclone rain
exposures tended to extended further inland compared to wind exposures, up to
the Appalachian mountains. This agrees with previous research indicating that
the Appalachian mountains' topography both enhances precipitation during
tropical cyclones and provides hydrological conditions for severe
flooding~\parencite{rees2001}.  Almost all tropical cyclone tornado exposures
were in southern coastal states, consistent with previous evidence that most
noteworthy tropical cyclone-related tornadoes occur on the right-side of the
tropical cyclone track in Atlantic-basin \ac{US} storms~\parencite{moore2012}.
The exposure averages we calculated may be limited as estimates of
long-term frequencies, as tropical cyclones follow decadal
patterns~\parencite{kossin2007more} likely not adequately captured in the
available data, which covered less than 20 years.

@@

We found that tropical cyclones tended to bring different hazards to different
counties, and so agreement was typically low between distance-based
tropical cyclone exposure assessment and each of the single hazard-specific
exposures, as well as between pairs of hazard-specific metrics
(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard} and S4).  These
findings align with previous results from atmospheric science and related
fields on the characteristics of tropical cyclones. While tropical cyclone
rainfall and windspeed can be well-correlated when the tropical cyclone is over
water~\parencite{cerveny2000}, this relationship often weakens once the
hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
cyclones bring higher risks of dangerous winds inland~\parencite{kruk2010},
while slow-moving tropical cyclones are likely to bring more
rain~\parencite{rappaport2000} and cause more damage because of sustained
hazardous conditions~\parencite{rezapour2014}. Further, while the likelihood
and extent of flooding during a tropical cyclone is related to the tropical
cyclone's rainfall, it is also driven by factors like top soil saturation and
the structure of the water basin's drainage network~\parencite{chen2015,
rees2001}.

Based on our results, the use of a distance-based metric to assess exposure to any of
these hazards, or the use of measurements from one hazard as a proxy for
exposure to any of the other hazards considered, would often introduce exposure
misclassification. This conclusion reinforces similar findings from a study of
Florida's 2004 storm season~\parencite{grabich2015measuring}.  For some
studies, such exposure misclassification might plausibly be differential.  For
example, tropical cyclone wind exposures tend to be concentrated in counties
near the coast, while tropical cyclone wind exposures sometimes extend well
inland.  If the etiologically-relevant exposure for a health outcome is extreme
rainfall but exposure is classified based on measurements of wind, the
probability of being misclassified as unexposed would be higher in inland
counties, while the probability of being misclassified as exposed would be
higher in coastal counties. If coastal counties differ from inland counties in
either the outcome of interest or in factors associated with risk of that
outcome, exposure misclassification would be
differential~\parencite{savitz2016interpreting}.  Such differential exposure
misclassification could bias estimates of tropical cyclone effects either
towards the null (estimating a lower or null association compared to the true
association that exists) or away from the null (estimating a larger association
than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.

We did, however, find a small set of tropical cyclones for which for which
agreement was high across several single-hazard exposure assessments (e.g.,
Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto in~2006
(Figure~\ref{fig:jaccard})).  Hurricanes Floyd in~1999 and Irene in~2011 both
made their first \ac{US} landfall in North Carolina at minor hurricane
windspeeds (Category~2 and~1, respectively) and then skimmed the eastern coast
of the \ac{US} north through New England, bringing substantial rainfall to much
of the eastern coast from North Carolina north and causing extensive inland
flooding in North Carolina (Floyd) and New England
(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
in~2008, Bertha in~1996, and Ernesto in~2006 also followed the eastern
coastline. For these storms, the tropical cyclones' persistent proximity to
water may have helped maintain wind speeds in similar patterns to rain and
distance-based exposure assessments, resulting in more similarities across
exposure assessments compared to other tropical cyclones.  For these storms, it
may be possible to assess exposure to multiple hazards of the storm using a
single metric, in some cases even a proxy like the distance between the county
and the storm's track.  However, for these storms it may be difficult to
untangle the contribution of each hazard to the overall effect of the storm,
given that several hazards have similar geographical patterns.

@@

The dataset presented here does have several limitations, in addition to the
caveats previously discussed. First, the dataset is not comprehensive of all
tropical cyclone hazards. For example, coastal counties can experience
dangerous storm surge, which is not specifically covered in this dataset
(although some resulting coastal flooding is captured). We are exploring ways
to include this in future versions of the dataset, but to date we have
focused on exposures that could affect any county, whether inland or coastal.
Second, this data is aggregated to the county level. Many health outcome
datasets are aggregated at this level, but some may be aggregated at a finer
spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e.,
point locations for each outcome). We have published the wind model used to
create this dataset as its own open-source R package
\parencite{stormwindmodel}, and it can be used to model storm-associated winds
at a finer spatial resolution; however, measurements of other hazards cannot
similarly be re-scaled through tools we provide. Next, we provide these data
and associated software tools through R packages, and so some experience in the
R programming language is required to make full use of them. However, R is
currently a popular programming language for environmental epidemiology,
allowing the data to reach a large audience, and we are exploring options to
create a web application using the Shiny platform to allow broader web-based
access of the data \parencite{shiny2019}.  Finally, to assess patterns and
agreement for binary exposure classifications, we have chosen one set of
sensible thresholds for binary classifications based on continuous metrics
(rainfall, maximum sustained surface wind, and distance from the storm's track).
Results and conclusions would differ somewhat with other threshold choices. We
have published code for this analysis online, allowing
other researchers to explore other threshold choices for these
analyses.

>>>>

Epidemiologic studies can help characterize which health risks are elevated
during disasters, to what degree, and for
whom~\parencite{ibrahim2005unfortunate, noji2005disasters}.  As a result, these
studies help improve disaster preparedness and
response~\parencite{noji2005disasters}.  However, tropical cyclones are
multi-hazard events, making it complicated to measure exposure and so to
conduct multi-community, multi-year studies leveraging large administrative
datasets.  Here, we provide open-source county-level data for several tropical
cyclone exposures and explore limitations in that data.  
Further, we explore patterns in storm exposure classifications based on
different metrics, and we find that county-level tropical cyclone exposure
assessments vary substantially when using different metrics.
Our results can inform exposure assessment for future county-level studies of
the health risk and impacts associated with tropical cyclones exposure as
well as provide insights to inform epidemiologic study design.

@@

The open-source data generally correspond well with data from other potential
sources (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison}
and~S2), but there are some caveats. The rainfall data are generally
well-correlated with ground-based observations, but may sometimes underestimate
very high rainfall values (Figure~\ref{fig:raincomparison}). When rainfall data
is used to create binary exposure classifications, this disagreement is
unlikely to influence results, as both data sources agree in identifying these
as storms with high rainfall, but would be important to consider for cases that
include rainfall as a continuous measurement.

The peak sustained wind estimates are based on modeled, rather than observed,
values, and while the modeled wind data generally agree well with post-analysis
maximum wind radii \parencite{landsea2013}
(Figure~\ref{fig:windcomparison}), there were a few storms with some
discrepancies. These storms---for example, Hurricane Sandy in 2012 and
Hurricane Ike in 2008---were unusually large systems for which high winds
persisted well inland from landfall (Figures~\ref{fig:windcomparison} and~S2).

For the flooding data, we found that flood event status, as determined based on
the NOAA Storm Events listings, typically agreed with measurements from
\ac{USGS} streamgages, with a flood event more likely to be listed if a storm
elevated streamflow at streamgages across the county
(Figure~\ref{fig:floodcomparison}).  However, there are differences between the
two flooding datasets, and these highlight both the difficulty of measuring
flood exposure at the county level and inherent challenges in using data from a
storm event database for epidemiologic exposure assessment.  For example, there
was one storm in Fulton County, GA, for which there was high streamflow but not
an associated flood event listing (Hurricane Ida, 2009).  This storm occurred
in November 2009, following a month with historic rainfall and flooding in
Georgia~\parencite{shepherd2011overview}.  In this case, the flooding
associated with Ida was incorporated into an ongoing flood event listing, with
a start date well before the five-day window we used to temporally match storm
event listings with tropical cyclone tracks for our data.

This disagreement highlights the difficulty of large-scale pairing of storm
tracks with storm event listings---without criteria for temporally matching
event start dates to storm dates, many false positives would be
captured, for which the occurrance of a storm in the midst of an ongoing storm
event might be improperly attributed to the storm.  However, distance
and time restrictions, like those we used in matching NOAA Storm Event listings
with tropical cyclone locations and dates, do cause 
occasional false negatives, as for Hurricane Ida in Fulton County, GA, where a
storm contributes meaningfully to an ongoing event, but the event is not
captured for the storm in the exposure data.

@@

Tropical cyclone wind exposures had a strong coastal pattern, which is
consistent with the dramatic decrease in wind intensity that typically
characterizes the landfall of tropical cyclones.  Tropical cyclone rain
exposures tended to extended further inland compared to wind exposures, up to
the Appalachian mountains. This agrees with previous research indicating that
the Appalachian mountains' topography both enhances precipitation during
tropical cyclones and provides hydrological conditions for severe
flooding~\parencite{rees2001}.  Almost all tropical cyclone tornado exposures
were in southern coastal states, consistent with previous evidence that
tropical cyclone-related tornadoes typically occur to the right of tropical
cyclone tracks in Atlantic-basin \ac{US} storms~\parencite{moore2012}.  It is
important to note, however, that the exposure averages we calculated may be
limited as estimates of long-term frequencies, as tropical cyclones follow
decadal patterns~\parencite{kossin2007more} that may not adequately captured in
the available data.

@@

We found that tropical cyclones tended to bring different hazards to different
counties, and so agreement was typically low between distance-based
tropical cyclone exposure assessment and each of the single hazard-specific
exposures, as well as between pairs of hazard-specific metrics
(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard} and S4).  These
findings align with previous results from atmospheric science and related
fields on the characteristics of tropical cyclones. While tropical cyclone
rainfall and windspeed can be well-correlated when the tropical cyclone is over
water~\parencite{cerveny2000}, this relationship often weakens once the
hurricane has made landfall~\parencite{jiang2008}.  Fast-moving tropical
cyclones bring higher risks of dangerous winds inland~\parencite{kruk2010},
while slow-moving tropical cyclones are likely to bring more
rain~\parencite{rappaport2000} and may cause more damage because of sustained
hazardous conditions~\parencite{rezapour2014}. Further, while the likelihood
and extent of flooding during a tropical cyclone is related to the tropical
cyclone's rainfall, it is also driven by factors like top soil saturation and
the structure of the water basin's drainage network~\parencite{chen2015,
rees2001}.

Based on our results, the use of a distance-based metric to assess exposure to any of
these hazards, or the use of measurements from one hazard as a proxy for
exposure to any of the other hazards considered, would often introduce exposure
misclassification. This conclusion reinforces similar findings from a study of
Florida's 2004 storm season~\parencite{grabich2015measuring}.  For some
studies, such exposure misclassification might plausibly be differential.  For
example, tropical cyclone wind exposures tend to be concentrated in counties
near the coast, while tropical cyclone rain exposures sometimes extend well
inland.  If the etiologically-relevant exposure for a health outcome is extreme
rainfall but exposure is classified based on measurements of wind, the
probability of being misclassified as unexposed would be higher in inland
counties, while the probability of being misclassified as exposed would be
higher in coastal counties. If coastal counties differ from inland counties in
either the outcome of interest or in factors associated with risk of that
outcome, exposure misclassification would be
differential~\parencite{savitz2016interpreting}.  Such differential exposure
misclassification could bias estimates of tropical cyclone effects either
towards the null (estimating a lower or null association compared to the true
association that exists) or away from the null (estimating a larger association
than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.

We did find a small set of tropical cyclones for which for which
agreement was high across several single-hazard exposure assessments (e.g.,
Floyd in~1999, Irene in 2011, Hannah in~2008, Bertha in~1996; Ernesto in~2006
(Figure~\ref{fig:jaccard})).  Hurricanes Floyd in~1999 and Irene in~2011 both
made their first \ac{US} landfall in North Carolina at minor hurricane
windspeeds (Category~2 and~1, respectively) and then skimmed the eastern coast
of the \ac{US} north through New England, bringing substantial rainfall to much
of the eastern coast from North Carolina north and causing extensive inland
flooding in North Carolina (Floyd) and New England
(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.  Hurricanes Hannah
in~2008, Bertha in~1996, and Ernesto in~2006 also followed the eastern
coastline. For these storms, the tropical cyclones' persistent proximity to
water may have helped maintain wind speeds in similar patterns to rain and
distance-based exposures, resulting in more similarities across
exposure assessments compared to other tropical cyclones.  For these storms, it
may be possible to assess exposure to multiple hazards of the storm using a
single metric, perhaps even a proxy like the distance between the county
and the storm's track.  However, for these storms it may be difficult to
untangle the contribution of each hazard to the overall effect of the storm,
given that several hazards have similar geographical patterns.

@@

The dataset presented here does have several limitations, in addition to the
caveats previously discussed. First, the dataset is not comprehensive of all
tropical cyclone hazards. For example, coastal counties can experience
dangerous storm surge, which is not specifically covered in this dataset
(although some resulting coastal flooding is captured). We are exploring ways
to include this in future versions of the dataset, but to date we have
focused on exposures that could affect any county, whether inland or coastal.
Second, these data are aggregated to the county level. Many health outcome
datasets are aggregated at this level, but some may be aggregated at a finer
spatial resolution (e.g., Census tract or ZIP code) or unaggregated (i.e.,
point locations for each outcome). We have published the wind model used to
create this dataset as its own open-source R package
\parencite{stormwindmodel}, and it can be used to model storm-associated winds
at a finer spatial resolution; however, measurements of other hazards cannot
similarly be re-scaled through tools we provide. Next, we provide these data
and associated software tools through R packages, and so some experience in the
R programming language is required to make full use of them. However, R is
currently a popular programming language for environmental epidemiology,
allowing the data to reach a large audience, and we are exploring options to
create a web application using the Shiny platform to allow broader web-based
access of the data \parencite{shiny2019}.  Finally, to assess patterns and
agreement for binary exposure classifications, we have chosen one set of
sensible thresholds for binary classifications based on continuous metrics
(rainfall, maximum sustained surface wind, and distance from the storm's track).
Results and conclusions would differ somewhat with other threshold choices. We
have published code for this analysis online, allowing
other researchers to explore other threshold choices for these
analyses.

### 2020-02-20 writing/ehp_introduction.tex
Edit the text

Hurricanes and other tropical cyclones can severely impact human health in
\ac{US} communities, as has been characterized in several long-term national
studies~\parencite{rappaport2000, rappaport2014fatalities,
rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.
These studies characterize tropical cyclone health impacts using fatality data
aggregated in large part from disaster-related mortality ascertainment and
surveillance conducted by agencies like the \ac{US} Centers for Disease Control
and Prevention, the \ac{US} National Weather Service, and local vital
statistics departments. These fatality data are based on case-by-case
ascertainment: identifying and characterizing deaths for which there is a clear
link with a disaster, through death certificate coding or other indicators
\parencite{rocha2017medicolegal}.

For several types of climate-related disasters---including heat waves, floods,
and wildfires---such research has been richly supplemented by studies that
estimate community-wide excess mortality and morbidity associated with the
disaster (e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
liu2017wildfire, milojevic2017mental}).  These studies, which often use time
series or case-crossover study designs, typically compare the rate of a health
outcome in a community during a disaster to rates during comparable
non-disaster periods.  While such studies cannot attribute specific cases
(e.g., a specific death) to a disaster, they can quantify the community-wide
change in health risk during disasters and capture impacts that might be missed
or underestimated with traditional disaster-related mortality ascertainment and
surveillance.  In many cases, these studies analyze multi-year, multi-community
data, allowing them to estimate average associations over many disasters and to
explore how a disaster's characteristics, or the characteristics of the
community it hits, modify associated health risks
(e.g.,~\cite{anderson2010heat, son2012impact, liu2017wildfire}).  Some studies
have begun to use this approach to study the health impacts of tropical
cyclones, including several studies of Hurricane Maria
(e.g.,~\cite{santos2018use, santos2018differential}), Hurricane Sandy
(e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane season
in Florida~\parencite{mckinney2011}.  However, to expand this approach to
longer time periods and large sets of communities, researchers must be able to
assess exposure to tropical cyclones consistently and comparably across storms,
years, and communities.

The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
standard for Atlantic-basin tropical cyclone tracks.  It records the central
position of a tropical cyclone every six hours, as well as the storm's minimum
central pressure and maximum sustained surface wind.  These data are openly
available through the \ac{HURDAT2}, a post-storm assessment that incorporates
data from several sources, including satellite data and, when available,
aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these
data, it is straightforward to measure whether a community was in a storm's
direct path or within a certain distance of its central track; some previous
epidemiological studies have done this to assess exposure to a tropical storm
(e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}).

However, when epidemiological studies assess exposure to a tropical cyclone
using, as a proxy, how close the storm came to a community, they
may misclassify exposure~\parencite{grabich2015measuring}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and
waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
floods, and tornadoes can occur well away from the tropical cyclone's central
track~\parencite{rappaport2000, atallah2007, moore2012}.  For example, fatal
tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
tropical cyclone's center~\parencite{moore2012}.  Further, when studies use
distance from the storm's track to assess exposure, they often use the same
buffer constraints on each side of the storm track
(e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
cyclone tend to be distributed around its center non-symmetrically.  Extreme
winds are more common to the track's right, where counter-clockwise cyclonic
winds move in concert with the tropical cyclone's forward
motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
is often heaviest to the left of the track, especially when the tropical
cyclone interacts with other weather systems~\parencite{atallah2003,
atallah2007, zhu2013variations} or undergoes an extratropical
transition~\parencite{elsberry2002}.  The multi-hazard nature of tropical
cyclones therefore makes it hard to assess exposure based on how close the
storm's central track came to the community.  While other approaches have been
developed to incorporate storm hazards, particulary wind, into exposure
assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
there is not yet a standard approach, and when different studies use different
datasets or storm hazards when assessing storm exposure, it becomes difficult
to compare and aggregate findings.

To help epidemiological researchers assess exposure to tropical cyclones, we
developed an open-source dataset~\parencite{hurricaneexposure}, which we
present here.  These data cover all counties in states in the eastern half of
the \ac{US} over multiple years and include five metrics characterizing
exposure to tropical cyclones (Table~\ref{tab:exposuremetrics}): (1)~closest
distance the storm's central track came to the county's center (a proxy
measurement of storm exposure used in some previous studies); (2)~peak
sustained surface wind at the county's center over the course of the storm;
(3)~cumulative rainfall in the county over the course of the storm;
(4)~flooding in the county concurrent with the storm; and (5)~tornadoes in the
county concurrent with the storm.  We aggregated these data at the county level
since data for epidemiological studies are often available at county-level
aggregations (e.g., direct hurricane-related deaths~\parencite{czajkowski2011};
birth outcomes~\parencite{grabich2015, grabich2015measuring}; autism
prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
for, and respond to, tropical cyclones are often undertaken at the county
level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
explore and map these data and to link it with county-level health
data~\parencite{hurricaneexposuredata}.

Here, we describe how we developed this dataset and explore how its
measurements compare to other data that could be used to characterize tropical
cyclone hazards for epidemiological research. Further, we expand on previous
research on methods to measure exposure to tropical cyclones for
epidemiological research~\parencite{grabich2015measuring}.  The previous study
investigated differences in which communities were assessed as exposed to
tropical cyclones during the 2004 hurricane season in Florida, depending on
whether exposure assessment was based on distance to the storm's track or a
metric incorporating storm-related winds within the county during the storm.
They found important differences across methods of assessing storm exposure,
concluding that a study may be prone to bias from exposure misclassification if
distance to the storm track is used as a proxy measurement for exposure
assessment~\parencite{grabich2015measuring}.  Here we expand to investigate
this question across a large set of counties and storm seasons.  Further, we
investigate patterns in exposure classification based on other metrics of
storm-related hazards---rainfall, flooding, and tornadoes---which are important
for inland health impacts of tropical cyclones~\parencite{czajkowski2011,
moore2012}. These results can help epidemiologists design studies and plan
statistical analysis for multi-year, multi-community studies that estimate
excess mortality and morbidity exposure to the hazards brought by tropical
cyclones.

>>>>

Hurricanes and other tropical cyclones can severely impact human health in
\ac{US} communities, as shown in several long-term national
studies~\parencite{rappaport2000, rappaport2014fatalities,
rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.
These studies characterize tropical cyclone health impacts using fatality data
aggregated in large part from disaster-related mortality ascertainment and
surveillance conducted by agencies like the \ac{US} Centers for Disease Control
and Prevention, the \ac{US} National Weather Service, and local vital
statistics departments. These fatality data are based on case-by-case
ascertainment: identifying and characterizing deaths for which there is a clear
link with a disaster, through death certificate coding or other indicators
\parencite{rocha2017medicolegal}.

For several types of climate-related disasters---including heat waves, floods,
and wildfires---such research has been richly supplemented by studies that
estimate community-wide excess mortality and morbidity associated with the
disaster (e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
liu2017wildfire, milojevic2017mental}).  These studies, which typically use time
series or case-crossover designs, compare the community-wide rate of a health
outcome during a disaster to rates during comparable
non-disaster periods.  While such studies cannot attribute specific cases
(e.g., a specific death) to a disaster, they can quantify the community-wide
change in health risk during disasters and capture impacts that might be missed
or underestimated with traditional disaster-related mortality ascertainment and
surveillance.  In many cases, these studies analyze multi-year, multi-community
data, allowing them to estimate average associations over many disasters and to
explore how a disaster's characteristics, or the characteristics of the
community it hits, modify associated health risks
(e.g.,~\cite{anderson2010heat, son2012impact, liu2017wildfire}).  Some studies
have begun to use this approach to study the health impacts of tropical
cyclones, including several studies of Hurricane Maria
(e.g.,~\cite{santos2018use, santos2018differential}), Hurricane Sandy
(e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane season
in Florida~\parencite{mckinney2011}.  However, to expand this approach to
longer time periods and larger sets of communities, researchers must be able to
assess exposure to tropical cyclones consistently and comparably across storms,
years, and communities.

The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
standard for Atlantic-basin tropical cyclone tracks.  It records the central
position of a tropical cyclone every six hours, as well as the storm's minimum
central pressure and maximum sustained surface wind.  These data are openly
available through the \ac{HURDAT2}, a post-storm assessment that incorporates
data from several sources, including satellite data and, when available,
aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these
data, it is straightforward to measure a storm's direct path, and so to measure
whether a community was on or within a certain distance of that central track.
Some previous epidemiological studies have done this to assess exposure to a
tropical storm (e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}), 
and this method can be applied consistently across years and communities 
for large-scale studies.

However, when epidemiological studies assess exposure to a tropical cyclone
using, as a proxy, how close the storm came to a community, they may
misclassify exposure~\parencite{grabich2015measuring}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and
waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
floods, and tornadoes can occur well away from the tropical cyclone's central
track~\parencite{rappaport2000, atallah2007, moore2012}.  For example,
tornadoes generated by tropical cyclones, which were linked to over~300 deaths
in the \ac{US} between~1995 and~2009, most often
occur~200\,--\,500~\si{\kilo\metre} from the tropical cyclone's
center~\parencite{moore2012}.  Further, when studies use distance from the
storm's track to assess exposure, they often use the same buffer constraints on
each side of the storm track (e.g.,~\cite{kinney2008, currie2013}).  However,
the forces of a tropical cyclone tend to be distributed around its center
non-symmetrically.  Extreme winds are more common to the track's right, where
counter-clockwise cyclonic winds move in concert with the tropical cyclone's
forward motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
is often heaviest to the left of the track, especially when the tropical
cyclone interacts with other weather systems~\parencite{atallah2003,
atallah2007, zhu2013variations} or undergoes an extratropical
transition~\parencite{elsberry2002}.  The multi-hazard nature of tropical
cyclones therefore makes it hard to assess exposure based on how close the
storm's central track came to the community.  While other approaches have been
developed to incorporate storm hazards, particulary wind, into exposure
assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
there is not yet a standard approach, and when different studies use
different datasets or storm hazards when assessing storm exposure, it becomes
difficult to compare and aggregate findings.

To help epidemiological researchers assess exposure to tropical cyclones, we
developed an open-source dataset~\parencite{hurricaneexposure}, which we
present here.  These data cover all counties in states in the eastern half of
the \ac{US} over multiple years and include five metrics characterizing
exposure to tropical cyclones (Table~\ref{tab:exposuremetrics}): (1)~closest
distance the storm's central track came to the county's center (a proxy
measurement of storm exposure used in some previous studies); (2)~peak
sustained surface wind at the county's center over the course of the storm;
(3)~cumulative rainfall in the county over the course of the storm;
(4)~flooding in the county concurrent with the storm; and (5)~tornadoes in the
county concurrent with the storm.  We aggregated these data at the county level
since data for epidemiological studies are often available at this level (e.g.,
direct hurricane-related deaths~\parencite{czajkowski2011}; birth
outcomes~\parencite{grabich2015, grabich2015measuring}; autism
prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
for, and respond to, tropical cyclones are often undertaken at the county
level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
explore and map these data and to link them with county-level health
data~\parencite{hurricaneexposuredata}.

Here, we describe how we developed this dataset and explore how its
measurements compare to other data that could be used to characterize tropical
cyclone hazards for epidemiologic research. Further, we expand on previous
research on methods to measure exposure to tropical cyclones for epidemiologic
research~\parencite{grabich2015measuring}. This previous study investigated
differences in which communities were assessed as exposed to tropical cyclones
during the 2004 hurricane season in Florida, comparing exposure assessment
based on distance to the storm's track or versus assessment based on a metric
that incorporated storm-generated winds within the county.  They found
important differences across methods of assessing storm exposure, concluding
that a study may be prone to bias from exposure misclassification if distance
to the storm track is used as a proxy measurement for exposure
assessment~\parencite{grabich2015measuring}.  Here we expand to investigate
this question across a larger set of counties and storm seasons.  Further, we
investigate patterns in exposure classification based on other metrics of
storm-related hazards---rainfall, flooding, and tornadoes---all of which are
important for inland health impacts of tropical
cyclones~\parencite{czajkowski2011, moore2012}. These results can help
epidemiologists design studies and plan statistical analysis for multi-year,
multi-community studies that estimate excess mortality and morbidity associated
with exposure to the hazards brought by tropical cyclones.
